HydroPowerModels.jl: A Julia/JuMP Package for Hydrothermal Economic Dispatch Optimization

### 2019-08-05 paper/paper.tex
update paper

% <review @Lukrosz>
% I was taught not to use apostrophes in articles. However, I have seen other articles in different journals where apostrophes are used. Therfore, I encourage you not to use them but the decision is up to you.
% </review>
A number of programs implementing SDDP are in-use around the world, ranging from unpublished implementations in academic institutions to professional software such as the product developed by PSR, a software and consulting company, also called \textit{SDDP}\footnote{Note that the ``SDDP'' acronym is used to denote a software product, the original algorithm of \cite{pereira1991multi}, and a more general class of algorithms inspired by the original SDDP algorithm.} \cite{psr_software_2019}. However, until recently, there was no fast, reliable, and open-source implementation of the SDDP algorithm. Without such a tool, researchers and practitioners have not had a common ground for the discussion and analysis of different hydrothermal dispatch formulations and their solutions.

@@

The next sections are organized as follows: 1) A brief understanding of multistage stochastic problems and decisions under uncertainty; 2) A presentation of the problem's modeling formulation and specification; 3) Different simplifications of the model; 4) An explanation of the solution method (SDDP); 5) A comparison of existing alternative solution implementations and listing of the proposed package functionalities; 6) A case study to clarify usage.

@@

Stochastic Programming (SP) \cite{shapiro2009lectures} is a branch of optimization under uncertainty, where the realization of some random variables ($\omega$) influence the problem's conditions and consequently the optimal decisions. Uncertainty relates to the probability distributions of parameters and may be incorporated to the problem in various manners.

@@

@@

\subsubsection*{Variables}
\begin{align*}
%
& S^g_k \;\; \forall k\in G \mbox{ - generator complex power dispatch} \\
& V_i \;\; \forall i\in N \mbox{ - bus complex voltage}\\
& S_{ij} \;\; \forall (i,j) \in E \cup E^R \mbox{ - branch complex power flow}\\
& U_h \;\; \forall h \in H \mbox{ - reservoir outflow}\\
& \mathfrak{S}_h \;\; \forall h \in H \mbox{ - reservoir spillage}\\
& \nu_h \;\; \forall h \in H \mbox{ - reservoir volume}\\
& D_i \;\; \forall i\in N \mbox{ - bus complex deficit}\\
%
\end{align*}

A complete mathematical model of the sub-problem is as follows,
\begin{subequations} \label{AC_OPF}
\begin{align}
%
& \mathcal{Q}_t(\nu_{t-1},\omega_t) = \nonumber \\
%
% \mbox{variables: } & \nonumber \\
% & S^g_k \;\; \forall k\in G \mbox{ - generator complex power dispatch} \label{var_generation}\\
% & V_i \;\; \forall i\in N \label{var_voltage} \mbox{ - bus complex voltage}\\
% & S_{ij} \;\; \forall (i,j) \in E \cup E^R  \label{var_complex_power} \mbox{ - branch complex power flow}\\
% & U_h \;\; \forall h \in H  \label{var_outflow} \mbox{ - reservoir outflow}\\
% & \mathfrak{S}_h \;\; \forall h \in H  \label{var_spillage} \mbox{ - reservoir spillage}\\
% & \nu_h \;\; \forall h \in H  \label{var_volume} \mbox{ - reservoir volume}\\
% & D_i \;\; \forall i\in N \label{var_deficit} \mbox{ - bus complex deficit}\\
%
\mbox{minimize: } & \sum_{k \in G} c_{2k} (\Re(S^g_k))^2 + c_{1k}\Re(S^g_k) + c_{0k} + \nonumber\\
& \qquad \sum_{i \in N} c_{d} \Re(D_i) +  \sum_{h \in H} c_{s} \mathfrak{S}_h + \rho_{t+1}[\mathcal{Q}_{t+1}(\nu_{t}, \omega_{t+1})] \label{eq_objective}\\
%
\mbox{subject to: } & \nonumber \\
& \angle V_{r} = 0  \;\; \forall r \in R \label{eq_ref_bus}\\
& D_i+ \sum_{\substack{k \in G_i}} S^g_k - \sum_{\substack{k \in L_i}} S^d_k - \sum_{\substack{k \in S_i}} Y^s_k |V_i|^2 \nonumber \\ 
& \qquad = \sum_{\substack{(i,j)\in E_i \cup E_i^R}} S_{ij} \;\; \forall i\in N \label{eq_kcl_shunt} \\
& S_{ij} = \left( Y_{ij} + Y^c_{ij}\right)^* \frac{|V_i|^2}{|{T}_{ij}|^2} - Y^*_{ij} \frac{V_i V^*_j}{{T}_{ij}} \;\; \forall (i,j)\in E \label{eq_power_from}\\
& S_{ji} = \left( Y_{ij} + Y^c_{ji} \right)^* |V_j|^2 - Y^*_{ij} \frac{V^*_i V_j}{{T}^*_{ij}} \;\; \forall (i,j)\in E \label{eq_power_to}\\
& |S_{ij}| \leq s^u_{ij} \;\; \forall (i,j) \in E \cup E^R \label{eq_thermal_limit}\\
& \theta^{\Delta l}_{ij} \leq \angle (V_i V^*_j) \leq \theta^{\Delta u}_{ij} \;\; \forall (i,j) \in E \label{eq_angle_difference}  \\
& \nu_h + U_h + \mathfrak{S}_h = \nu_{h,t-1} + a_h(\omega_t) + \nonumber \\ 
& \qquad  \sum_{\substack{k \in H^U_h}} U_k + \sum_{\substack{k \in H^{\mathfrak{S}}_h}} \mathfrak{S}_k \;\; \forall h \in H \label{eq_hydro_balance} \\
& U_h p_h = \Re(S^g_{G^H_h}) \;\; \forall h \in H^G \label{eq_hydro_generation} \\
& S^{gl}_k \leq S^g_k \leq S^{gu}_k \;\; \forall k \in G  \label{eq_gen_bounds}\\
& v^l_i \leq |V_i| \leq v^u_i \;\; \forall i \in N \label{eq_voltage_bounds}\\
& 0 \leq \nu_h \leq \overline{\nu}_h \;\; \forall h \in H \label{eq_volume_limit}\\
& 0 \leq U_h \leq \overline{u}_h \;\; \forall h \in H \label{eq_outflow_limit}
%
\end{align}
\end{subequations}
\footnote{Note that the elements of E are pairs of elements of N, so, in equations (\ref{eq_power_from}) and (\ref{eq_power_to}), variables defined over N are indexed by elements of E.}

The objective of the sub-problem (\ref{eq_objective}) is to minimize the costs of real power generation, cost of real energy supply's deficit, cost of spillage (in order to try and avoid degenerate solutions) and the cost-to-go function $\mathcal{Q}$.

@@

The AC Optimal Power Flow problem, defined in (\ref{AC_OPF}), captures the nonlinear and complex nature of the power flow. However, solving a nonlinear problem (NLP) is hard. The scarcity of efficient NLP solvers, the numerical issues created by large problem instances and the inconsistency of solutions (local optimums) reduce the depth of research using AC-OPF. Besides, the requirement of convexity frequently needed for applications limit its usage.

Hence, many approximations and relaxations have been developed for the AC-OPF \cite{molzahn2019survey}. In general these are simplifications and are easier to solve, but they ignore some parts of the problem. As a result, it is imperative to understand and weigh the advantages and compromises of each formulation when choosing one to use in a model.

@@

The \textit{SDP relaxation} deals with the non-convex constraints of the problem by using the fact that they define a positive semi-definite matrix with rank 1. The relaxation comes from removing the rank 1 restriction. This formulation's feasible region is contained within the SOC relaxation's feasible region, providing a better bound to the original problem.

@@

For this study, a 160 stage horizon planning and simulation have been executed using the following case parameters:

\begin{itemize}
    \item Number of stages: 160
    \item Number of hours in stage: 168 
    \item Network Formulation: Linear DC approximation
\end{itemize}

@@

Similar studies are possible for any case and formulation chosen, helping to analyze existing realistic cases and assess impacts of future system changes.

% \begin{figure}[!ht]\label{fig:thermal}
% \centering
% \includegraphics[width=0.5\columnwidth]{fig/plot_agg_result_thermal.png}
% \caption{Aggregated quantity of thermal generation.}
% \end{figure}

\begin{figure}[H]
\centerline{\includegraphics[width=0.5\linewidth]{fig/plot_agg_result_thermal.png}}
\label{fig:plot_agg_result_thermal}
\caption{Thermal Generation}
\end{figure}

\begin{figure}[H]
\centerline{\includegraphics[width=0.55\linewidth]{fig/plot_agg_result_load.png}}
\caption{Load Weighted Average Nodal Price}
\label{fig:plot_agg_result_load}
\end{figure}

\begin{figure}[H]
\centerline{\includegraphics[width=0.55\linewidth]{fig/plot_agg_result_def.png}}
\caption{Deficit}
\label{fig:plot_agg_result_def}
\end{figure}

\begin{figure}[H]
\centerline{\includegraphics[width=0.55\linewidth]{fig/plot_agg_result_hydrogen.png}}
\caption{Hydro Generation}
\label{fig:plot_agg_result_hydrogen}
\end{figure}

\begin{figure}[H]
\centerline{\includegraphics[width=0.55\linewidth]{fig/plot_agg_result_hydroturn.png}}
\caption{Hydro Turn (Storage Outflow)}
\label{fig:plot_agg_result_hydroturn}
\end{figure}

\begin{figure}[H]
\centerline{\includegraphics[width=0.5\linewidth]{fig/plot_agg_result_hydrospill.png}}
\caption{Hydro Spillage (Storage Spillage)}
\label{fig:plot_agg_result_hydrospill}
\end{figure}

\begin{figure}[H]
\centerline{\includegraphics[width=0.55\linewidth]{fig/plot_agg_result_vol.png}}
\caption{Volume Reservoir}
\label{fig:plot_agg_result_vol}
\end{figure}

\begin{figure}[H]
\centerline{\includegraphics[width=0.55\linewidth]{fig/plot_agg_result_inflow.png}}
\caption{Inflow}
\label{fig:plot_agg_result_inflow}
\end{figure}

>>>>

A number of programs implementing SDDP are in-use around the world, ranging from unpublished implementations in academic institutions to professional software such as the product developed by PSR, a software and consulting company, also called \textit{SDDP}\footnote{Note that the ``SDDP'' acronym is used to denote a software product, the original algorithm of \cite{pereira1991multi}, and a more general class of algorithms inspired by the original SDDP algorithm.} \cite{psr_software_2019}. However, until recently, there was no fast, reliable, and open-source implementation of the SDDP algorithm. Without such a tool, researchers and practitioners have not had a common ground for the discussion and analysis of different hydrothermal dispatch formulations and their solutions.

@@

The next sections are organized as follows: 1) A brief understanding of multistage stochastic problems and decisions under uncertainty; 2) A presentation of the modeling formulation and problem specification; 3) Different simplifications of the model; 4) An explanation of the solution method (SDDP); 5) A comparison of existing alternative solution implementations and listing of the proposed package functionalities; 6) A case study to clarify usage.

@@

Stochastic Programming (SP) \cite{shapiro2009lectures} is a branch of optimization under uncertainty, where the realization of some random variables ($\omega$) influence the conditions of the problem and consequently the optimal decisions. Uncertainty relates to the probability distributions of parameters and may be incorporated to the problem in various manners.

@@

\vspace{5mm}

@@

\subsubsection*{Variables}

\begin{align*}
%
& S^g_k \;\; \forall k\in G \mbox{ - generator complex power dispatch} \\
& V_i \;\; \forall i\in N \mbox{ - bus complex voltage}\\
& S_{ij} \;\; \forall (i,j) \in E \cup E^R \mbox{ - branch complex power flow}\\
& U_h \;\; \forall h \in H \mbox{ - reservoir outflow}\\
& \mathfrak{S}_h \;\; \forall h \in H \mbox{ - reservoir spillage}\\
& \nu_h \;\; \forall h \in H \mbox{ - reservoir volume}\\
& D_i \;\; \forall i\in N \mbox{ - bus complex deficit}\\
%
\end{align*}

A complete mathematical model of the sub-problem is as follows,
\begin{subequations} \label{AC_OPF}
\begin{align}
%
& \mathcal{Q}_t(\nu_{t-1},\omega_t) = \nonumber \\
%
\mbox{minimize: } & \sum_{k \in G} c_{2k} (\Re(S^g_k))^2 + c_{1k}\Re(S^g_k) + c_{0k} + \nonumber\\
& \qquad \sum_{i \in N} c_{d} \Re(D_i) +  \sum_{h \in H} c_{s} \mathfrak{S}_h + \rho_{t+1}[\mathcal{Q}_{t+1}(\nu_{t}, \omega_{t+1})] \label{eq_objective}\\
%
\mbox{subject to: } & \nonumber \\
& \angle V_{r} = 0  \;\; \forall r \in R \label{eq_ref_bus}\\
& D_i+ \sum_{\substack{k \in G_i}} S^g_k - \sum_{\substack{k \in L_i}} S^d_k - \sum_{\substack{k \in S_i}} Y^s_k |V_i|^2 \nonumber \\ 
& \qquad = \sum_{\substack{(i,j)\in E_i \cup E_i^R}} S_{ij} \;\; \forall i\in N \label{eq_kcl_shunt} \\
& S_{ij} = \left( Y_{ij} + Y^c_{ij}\right)^* \frac{|V_i|^2}{|{T}_{ij}|^2} - Y^*_{ij} \frac{V_i V^*_j}{{T}_{ij}} \;\; \forall (i,j)\in E \label{eq_power_from}\\
& S_{ji} = \left( Y_{ij} + Y^c_{ji} \right)^* |V_j|^2 - Y^*_{ij} \frac{V^*_i V_j}{{T}^*_{ij}} \;\; \forall (i,j)\in E \label{eq_power_to}\\
& |S_{ij}| \leq s^u_{ij} \;\; \forall (i,j) \in E \cup E^R \label{eq_thermal_limit}\\
& \theta^{\Delta l}_{ij} \leq \angle (V_i V^*_j) \leq \theta^{\Delta u}_{ij} \;\; \forall (i,j) \in E \label{eq_angle_difference}  \\
& \nu_h + U_h + \mathfrak{S}_h = \nu_{h,t-1} + a_h(\omega_t) + \nonumber \\ 
& \qquad  \sum_{\substack{k \in H^U_h}} U_k + \sum_{\substack{k \in H^{\mathfrak{S}}_h}} \mathfrak{S}_k \;\; \forall h \in H \label{eq_hydro_balance} \\
& U_h p_h = \Re(S^g_{G^H_h}) \;\; \forall h \in H^G \label{eq_hydro_generation} \\
& S^{gl}_k \leq S^g_k \leq S^{gu}_k \;\; \forall k \in G  \label{eq_gen_bounds}\\
& v^l_i \leq |V_i| \leq v^u_i \;\; \forall i \in N \label{eq_voltage_bounds}\\
& 0 \leq \nu_h \leq \overline{\nu}_h \;\; \forall h \in H \label{eq_volume_limit}\\
& 0 \leq U_h \leq \overline{u}_h \;\; \forall h \in H \label{eq_outflow_limit}
%
\end{align}
\end{subequations}
\footnote{Note that the elements of E are pairs of elements of N, so, in equations (\ref{eq_power_from}) and (\ref{eq_power_to}), variables defined over N are indexed by elements of E.}

The objective of the sub-problem (\ref{eq_objective}) is to minimize the costs of real power generation, cost of deficit from the real energy supply, cost of spillage (in order to try and avoid degenerate solutions) and the cost-to-go function $\mathcal{Q}$.

@@

The AC Optimal Power Flow problem, defined in (\ref{AC_OPF}), captures the nonlinear and complex nature of the power flow. However, solving a nonlinear problem (NLP) is hard. Guaranteeing global optimality often involves an extensive number of comparisons and, thus, it is not done. Local optimums are used instead, providing inconsistent solutions. Along with the scarcity of efficient NLP solvers and the numerical issues created by large problem instances, the depth of research using AC-OPF is reduced. Besides, the requirement of convexity frequently needed for applications limit its usage. 
Hence, many approximations and relaxations have been developed for the AC-OPF \cite{molzahn2019survey}. In general these are simplifications and are easier to solve, but they ignore some parts of the problem. As a result, it is imperative to understand and weigh the advantages and compromises of each formulation when choosing one to use in a model.

@@

The \textit{SDP relaxation} deals with the non-convex constraints of the problem by using the fact that they define a positive semi-definite matrix with rank 1. The relaxation comes from removing the rank 1 restriction. The feasible region of this formulation is contained within the feasible region of the SOC relaxation, providing a better bound to the original problem.

@@

For this study, a 52 stage horizon planning and simulation have been executed using the following case parameters:

\begin{itemize}
    \item Number of stages: 52
    \item Number of hours in stage: 168 
    \item Network Formulation: Transportation model relaxation
\end{itemize}

@@

Similar studies are possible for any case and formulation chosen, helping to analyze existing realistic cases and assess impacts of future system changes. For example, we reran the study using a SOC relaxation to compare results. Most quantities analyzed, as those presented in the aggregated results, are similar, but operation costs differ as it is observable in figures \ref{fig:hist_cost_tube},\ref{fig:hist_cost_soc}. This is expected, since the SOC relaxation is a more realistic representation of reality.

Moreover, it is important to point out this extra cost would be even bigger if we used the policy of the transportation model to operate the SOC model in the simulations. In other words, there is a hidden cost when planning a dispatch associated with the fact we are not evaluating costs on the actual system we want to operate.

Therefore, measuring the impacts of possible simplifications is a needed step in discussing hydrothermal economic dispatch. \textit{HydroPowerModels.jl} intends to provide a common ground for discussions and analysis and a easy to use tool for research and applications.

\begin{figure}[H]
\centerline{\includegraphics[width=0.7\linewidth]{fig/plot_agg_result_thermal.png}}
\label{fig:plot_agg_result_thermal}
\caption{Thermal Generation}
\end{figure}

\begin{figure}[H]
\centerline{\includegraphics[width=0.7\linewidth]{fig/plot_agg_result_load.png}}
\caption{Load Weighted Average Nodal Price}
\label{fig:plot_agg_result_load}
\end{figure}

\begin{figure}[H]
\centerline{\includegraphics[width=0.7\linewidth]{fig/plot_agg_result_def.png}}
\caption{Deficit}
\label{fig:plot_agg_result_def}
\end{figure}

\begin{figure}[H]
\centerline{\includegraphics[width=0.7\linewidth]{fig/plot_agg_result_hydrogen.png}}
\caption{Hydro Generation}
\label{fig:plot_agg_result_hydrogen}
\end{figure}

\begin{figure}[H]
\centerline{\includegraphics[width=0.7\linewidth]{fig/plot_agg_result_hydroturn.png}}
\caption{Reservoir Outflow}
\label{fig:plot_agg_result_hydroturn}
\end{figure}

\begin{figure}[H]
\centerline{\includegraphics[width=0.7\linewidth]{fig/plot_agg_result_hydrospill.png}}
\caption{Reservoir Spillage}
\label{fig:plot_agg_result_hydrospill}
\end{figure}

\begin{figure}[H]
\centerline{\includegraphics[width=0.7\linewidth]{fig/plot_agg_result_vol.png}}
\caption{Volume Reservoir}
\label{fig:plot_agg_result_vol}
\end{figure}

\begin{figure}[H]
\centerline{\includegraphics[width=0.7\linewidth]{fig/plot_agg_result_inflow.png}}
\caption{Inflow}
\label{fig:plot_agg_result_inflow}
\end{figure}

\begin{figure}[H]
\centerline{\includegraphics[width=0.9\linewidth]{fig/hist_cost_tube.png}}
\caption{Simulation Costs Transportation}
\label{fig:hist_cost_tube}
\end{figure}

\begin{figure}[H]
\centerline{\includegraphics[width=0.9\linewidth]{fig/hist_cost_soc.png}}
\caption{Simulation Costs SOC}
\label{fig:hist_cost_soc}
\end{figure}
